Visualize ukb whole genome, using ctwas results

### 2021-01-27 analysis/index.Rmd
ctwas para results

>>>>

## `ctwas` package
Test the summary statistics version of the `ctwas` algorithm. Now provided as a R package.

* [Using UKbiobank data (22k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s40.22.html)

* [Using UKbiobank data (22k samples), fusion weights Adipose, in-sample LD](simulation-ctwas-ukbWG-gtex.adipose_s40.22_2.html)

* [Using UKbiobank data (45k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s80.45.html)

* [Using UKbiobank data (110k samples), fusion weights Adipose](simulation-ctwas-ukbWG-gtex.adipose_s200.113.html)

  * [Visualization](simulation-ctwas-ukbWG-gtex.adipose_visual.html)

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22.Rmd
ctwas para results

---
title: "Simulations to test ctwas summary stats version, 113k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s200.113/ukb-s200.113.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210118/" # /
runtag = "ukb-s200.113-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

```{r parameters}
n.ori <- 200000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

@@

  * **LD genotype reference**
  We randomly selected 2000 samples to serve as the LD reference.

@@

```{r snp}
load("data/power_s200.113.Rd")
# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
print(p1)
```

@@

```{r gene}
# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
print(p2)
# save(p1,p2, file = "data/power_s200.113.Rd")
```

@@

plot_PIP <- function(configtag, tags){
   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")

@@

## ctwas_rss

@@

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, simutags)
# plot_PIP(configtag, simutags)
```

>>>>

---
title: "Simulations to test ctwas summary stats version, 22k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

```{r parameters}
n.ori <- 40000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

@@

  * **LD genotype reference**
  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.

@@

```{r snp}
load("data/power_s40.22.Rd")
# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
print(p1)
```

@@

```{r gene}
# p2 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_beta"]], gene.p)
print(p2)
# save(p1,p2, file = "data/power_s80.45.Rd")
```

@@

plot_PIP <- function(configtag, tags){
   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

@@

## ctwas_rss parameter estimation

@@

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, simutags)
# plot_PIP(configtag, simutags)

simutags <- paste(3, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(4, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(5, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(6, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(7, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(8, 1:5, sep = "-")
plot_par(configtag, simutags)
```

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s40.22_2.Rmd
ctwas para results

---
title: "Simulations to test ctwas summary stats version, 22k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

  * **LD genotype reference**
  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.

@@

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, simutags)
# plot_PIP(configtag, simutags)

simutags <- paste(3, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(4, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(5, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(6, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(7, 1:5, sep = "-")
plot_par(configtag, simutags)

simutags <- paste(8, 1:5, sep = "-")
plot_par(configtag, simutags)
```

>>>>

---
title: "Simulations to test ctwas summary stats version, 22k samples, in-sample LD"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210112/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

  * **LD genotype reference**
  We used the same genotype as in simulation to serve as the LD reference.

@@

simutags <- paste(2, 1:5, sep = "-")
plot_par(configtag, simutags)
# plot_PIP(configtag, simutags)
```

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_s80.45.Rmd
ctwas para results

---
title: "Simulations to test ctwas summary stats version, 22k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s40.22/ukb-s40.22.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210113/" # /
runtag = "ukb-s40.22-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

```{r parameters}
n.ori <- 40000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

@@

  * **LD genotype reference**
  We used 2k samples from UKbiobank with same ethnicity to serve as the LD reference.

@@

```{r snp}
load("data/power_s40.22.Rd")
# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
print(p1)
```

@@

## ctwas_rss parameter estimation

>>>>

---
title: "Simulations to test ctwas summary stats version, 45k samples"
output: html_document
---

@@

```{r files}
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210117/" # /
runtag = "ukb-s80.45-adi"
configtags = 1
simutags = paste(rep(1:2, each = length(1:5)), 1:5, sep = "-")

@@

```{r parameters}
n.ori <- 80000 # number of samples
n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes
```

@@

  * **LD genotype reference**
  We randomly selected 2000 samples to serve as the LD reference.

@@

```{r snp}
load("data/power_s80.45.Rd")
# p1 <- pow(niter, n, phenores[["batch"]][[1]][["sigma_theta"]], snp.p)
print(p1)
```

@@

## ctwas_rss

### 2021-01-27 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd
ctwas para results

---
title: "Visualize ukb whole genome, s40.22"
date: "`r Sys.Date()`"
output: html_document
---

@@

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"

source("analysis/visual_twas_plots.R")
```

@@

plot_all <- function(susief, pf, phenof, chrom, ldres = NULL){
  a <- fread(susief, header =T)
  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  b <- rbind(b1, b2)
  setnames(b, old = c('X.CHROM', 'PVALUE', 'MARKER_ID'),
         new = c("chr", 'pvalue', 'name'))

  load(phenof)

  a <- a[a$chr == chrom,]
  b <- b[b$chr == chrom,]

  a <- merge(a, b, by = "name", all.x =T)

@@

  a <- a[, ifcausal:=as.character(ifcausal)]
  a[a$ifcausal == "1", "ifcausal"] = "Causal"
  a[a$ifcausal == "0", "ifcausal"] = "Non causal"

@@

```{r sim 1-3-17, out.width = "98%", out.height="95%"}
chrom <- 2
tag <- "20201001-2-5"
susief <-  paste0(outputdir, tag, ".config16.flrerun.susieI.txt")
pf <- paste0(simdatadir, tag)
phenof <- paste0(simdatadir, "simu_", tag, "-pheno.Rd")
plot_all(susief, pf, phenof, chrom)
```

>>>>

---
title: "Visualize ukb whole genome, using ctwas results"
date: "`r Sys.Date()`"
output: html_document
---

@@

```{r output dir}
outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210112/"

source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/visual_twas_plots.R")
```

@@

plot_all <- function(susief, pf, phenof, chr, ldres = NULL, thin = 1){
  a <- fread(susief, header =T)
  setnames(a, old = "id", new = "name")
  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  setnames(b1, old = "pos", new = "p0")
  b <- rbind(b1, b2, fill = T)
  setnames(b, old = c('PVALUE', 'id'),
         new = c('pvalue', 'name'))

  load(phenof)
  cau <- get_causal_id(phenores)

  a <- a[a$chrom == chr,]
  b <- b[b$chrom == chr,]

  a <- merge(a, b, by = "name")
  a$ifcausal <- ifelse(a$name %in% cau, "Causal", "Non causal" )

@@

@@

```{r sim 1-3-17, out.width = "98%", out.height="95%"}
# TODO, thin option
chrom <- 1
runtag <- "ukb-s40.22-adi"
simutag <- "1-1"
configtag <- 1

susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrss.txt")
pf <- paste0(outputdir, runtag, "_simu",simutag)
phenof <- paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd")
plot_all(susief, pf, phenof, chrom)
```
